### 2024-03-05 paper/paper.qmd
innitial commit with the structure and necessary componenets

---
title: "US General Society Survey Analysis"
author: "Shreya Sakura Noskor"
thanks: "Code and data are available at: [repository](https://github.com/NotSakura/USSocialSurvey.git)"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-03-09 paper/paper.qmd
cleaned data available

@@
 thanks: "Code and data are available at: [repository](https://github.com/NotSakura/USSocialSurvey.git)"
+classoption: abstract
+author: "Cristina Burca, Yan Mezhiborsky, Sakura Noskor"
+format: pdf
 date: today
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
-format: pdf
-number-sections: true
+abstract: |
+ Some abstact
+include-in-header: 
+  text:
+    \renewcommand{\abstractname}{Abstract}
 bibliography: references.bib
+margins:
+  top: 0.25in
+  bottom: 0.25in
+  left: 0.5in
+  right: 0.5in
+toc: true
 ---
@@
 library(tidyverse)
-library(palmerpenguins)
+```
+
+```{r}
+library(dataverse)
+library(tidyverse)
+
+ces2020 <-
+  get_dataframe_by_name(
+    filename = "CES20_Common_OUTPUT_vv.csv",
+    dataset = "10.7910/DVN/E9N6PH",
+    server = "dataverse.harvard.edu",
+    .f = read_csv
+  ) |>
+  select(votereg, CC20_410, employ, immstat, dualcit, gender, birthyr, race, inputstate, presvote16post, religpew, edloan)
+
+write_csv(ces2020, "ces2020.csv")
+
+```
+
+```{r}
+ces2020 <-
+  read_csv(
+    "ces2020.csv",
+    col_types =
+      cols(
+        "votereg" = col_integer(),
+        "CC20_410" = col_integer(),
+        "employ" = col_integer(),
+        "dualcit" = col_integer(), 
+        "gender" = col_integer(), 
+        "birthyr" = col_integer(), 
+        "race" = col_integer(), 
+        "inputstate" = col_integer(), 
+        "presvote16post" = col_integer(), 
+        "religpew" = col_integer(), 
+        "edloan" = col_integer()
+      )
+  )
+
+ces2020
+```
+
+```{r}
+
+ces2020 <-
+  ces2020 |>
+  filter(votereg == 1,
+         CC20_410 %in% c(1, 2)) |>
+  mutate(
+    voted_for = if_else(CC20_410 == 1, "Biden", "Trump"),
+    voted_for = as_factor(voted_for),
+    gender = if_else(gender == 1, "Male", "Female"),
+    employ = case_when(
+      employ == 1 ~ "Full-time",
+      employ == 2 ~ "Part-time",
+      employ == 3 ~ "laid off",
+      employ == 4 ~ "unemployed",
+      employ == 5 ~ "retired",
+      employ == 6 ~ "permannetly disabled", 
+      employ == 7 ~ "Homemaker", 
+      employ == 8 ~ "Student", 
+      employ == 9 ~ "Other"
+    ),
+    employ = factor(
+      employ,
+      levels = c(
+        "Full-time",
+        "Part-time",
+        "laid off",
+        "unemployed",
+        "retired",
+        "permannetly disabled", 
+        "Homemaker", 
+        "Student", "Other"
+      )
+    ), 
+    immstat = case_when(
+      immstat == 1 ~ "immigrant and citizen",
+      immstat == 2 ~ "immigrant not citizen",
+      immstat == 3 ~ "born in US, but parent(s) immigrant",
+      immstat == 4 ~ "parent and I born in US but grandparent(s) immigrant",
+      immstat == 5 ~ "all born in US",
+    ),
+    immstat = factor(immstat, levels = c( "immigrant and citizen","immigrant not citizen", "born in US, but parent(s) immigrant",     "parent and I born in US but grandparent(s) immigrant", "all born in US")
+    ),
+    dualcit = if_else(dualcit == 1, "Yes", "No"),
+    race = case_when(
+      race == 1 ~ "White",
+      race == 2 ~ "Black",
+      race == 3 ~ "Hispanic",
+      race == 4 ~ "Asian",
+      race == 5 ~ "Native American",
+      race == 6 ~ "Middle Eastern", 
+      race == 7 ~ "Two or more races", 
+      race == 8 ~ "Other"
+    ),
+    race = factor(race, levels = c(
+        "White","Black","Hispanic", "Asian", "Native American", "Middle Eastern",  "Two or more races", "Other")
+    ), 
+    presvote16post = case_when(
+      presvote16post == 1 ~ "Hillary Clinton",
+      presvote16post == 2 ~ "Donald Trump",
+      presvote16post == 3 ~ "Gary Johnson",
+      presvote16post == 4 ~ "Jill Stein",
+      presvote16post == 5 ~ "Evan McMullin",
+      presvote16post == 6 ~ "Other", 
+      presvote16post == 7 ~ "Did not vote"
+    ),
+    presvote16post = factor(presvote16post, levels = c(
+        "Hillary Clinton","Donald Trump","Gary Johnson", "Jill Stein", "Evan McMullin", "Other", "Did not vote")
+    ),
+    religpew = case_when(
+      religpew == 1 ~ "Protestant",
+      religpew == 2 ~ "Roman Catholic",
+      religpew == 3 ~ "Mormon",
+      religpew == 4 ~ "Eastern or Greek Orthodox",
+      religpew == 5 ~ "Jewish",
+      religpew == 6 ~ "Muslim", 
+      religpew == 7 ~ "Buddhist", 
+      religpew == 8 ~ "Hindu", 
+      religpew == 9 ~ "Atheist", 
+      religpew == 10 ~ "Agnostic",
+      religpew == 11 ~ "Nothing in Particular",
+      religpew == 12 ~ "Something else"
+    ),
+    religpew = factor(
+      religpew,
+      levels = c(
+        "Protestant",
+        "Roman Catholic",
+        "Mormon",
+        "Eastern or Greek Orthodox",
+        "Jewish",
+        "Muslim", 
+        "Buddhist", 
+        "Hindu", "Atheist", "Agnostic", "Nothing in Particular", "Something else"
+      )
+    ), 
+    edloan = case_when(
+      edloan == 1 ~ "Yes",
+      edloan == 2 ~ "No",
+      edloan == 3 ~ "Other",
+      edloan == 4 ~ "Don't know",
+      edloan == 5 ~ "None of these",
+    ),
+    edloan = factor(edloan, levels = c( "Yes","No", "Other", "Don't Know", "None of these")
+    ) #########ADD input state it if you want to (sakura)
+  ) |>
+  select(votereg, employ, immstat, dualcit, gender, birthyr, race, inputstate, presvote16post, religpew, edloan)
+
 ```
@@
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
-
-```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
-
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
-
-Talk way more about it.
-
 # Model
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
 
@@
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
 ### Model justification
@@
 # Results
-
-Our results are summarized in @tbl-modelresults.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
-```
 

### 2024-03-09 paper/paper.qmd
included model and echo code

--- 2024-03-05 paper/paper.qmd
@@
 title: "US General Society Survey Analysis"
-author: "Shreya Sakura Noskor"
 thanks: "Code and data are available at: [repository](https://github.com/NotSakura/USSocialSurvey.git)"
+classoption: abstract
+author: "Cristina Burca, Yan Mezhiborsky, Sakura Noskor"
+format: pdf
 date: today
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
-format: pdf
-number-sections: true
+abstract: |
+ Some abstact
+include-in-header: 
+  text:
+    \renewcommand{\abstractname}{Abstract}
 bibliography: references.bib
+margins:
+  top: 0.25in
+  bottom: 0.25in
+  left: 0.5in
+  right: 0.5in
+toc: true
 ---
@@
 library(tidyverse)
-library(palmerpenguins)
+```
+
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+library(dataverse)
+library(tidyverse)
+
+ces2020 <-
+  get_dataframe_by_name(
+    filename = "CES20_Common_OUTPUT_vv.csv",
+    dataset = "10.7910/DVN/E9N6PH",
+    server = "dataverse.harvard.edu",
+    .f = read_csv
+  ) |>
+  select(votereg, CC20_410, employ, immstat, dualcit, gender, birthyr, race, inputstate, presvote16post, religpew, edloan)
+
+write_csv(ces2020, "ces2020.csv")
+
+```
+
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+ces2020 <-
+  read_csv(
+    "ces2020.csv",
+    col_types =
+      cols(
+        "votereg" = col_integer(),
+        "CC20_410" = col_integer(),
+        "employ" = col_integer(),
+        "dualcit" = col_integer(), 
+        "gender" = col_integer(), 
+        "birthyr" = col_integer(), 
+        "race" = col_integer(), 
+        "inputstate" = col_integer(), 
+        "presvote16post" = col_integer(), 
+        "religpew" = col_integer(), 
+        "edloan" = col_integer()
+      )
+  )
+
+```
+
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+ces2020 <-
+  ces2020 |>
+  filter(votereg == 1,
+         CC20_410 %in% c(1, 2)) |>
+  mutate(
+    voted_for = if_else(CC20_410 == 1, "Biden", "Trump"),
+    voted_for = as_factor(voted_for),
+    gender = if_else(gender == 1, "Male", "Female"),
+    employ = case_when(
+      employ == 1 ~ "Full-time",
+      employ == 2 ~ "Part-time",
+      employ == 3 ~ "Laid off",
+      employ == 4 ~ "Unemployed",
+      employ == 5 ~ "Retired",
+      employ == 6 ~ "Permannetly disabled", 
+      employ == 7 ~ "Homemaker", 
+      employ == 8 ~ "Student", 
+      employ == 9 ~ "Other"
+    ),
+    employ = factor(
+      employ,
+      levels = c(
+        "Full-time",
+        "Part-time",
+        "laid off",
+        "unemployed",
+        "retired",
+        "permannetly disabled", 
+        "Homemaker", 
+        "Student", "Other"
+      )
+    ), 
+    immstat = case_when(
+      immstat == 1 ~ "immigrant and citizen",
+      immstat == 2 ~ "immigrant not citizen",
+      immstat == 3 ~ "born in US, but parent(s) immigrant",
+      immstat == 4 ~ "parent and I born in US but grandparent(s) immigrant",
+      immstat == 5 ~ "all born in US",
+    ),
+    immstat = factor(immstat, levels = c( "immigrant and citizen","immigrant not citizen", "born in US, but parent(s) immigrant",     "parent and I born in US but grandparent(s) immigrant", "all born in US")
+    ),
+    dualcit = if_else(dualcit == 1, "Yes", "No"),
+    race = case_when(
+      race == 1 ~ "White",
+      race == 2 ~ "Black",
+      race == 3 ~ "Hispanic",
+      race == 4 ~ "Asian",
+      race == 5 ~ "Native American",
+      race == 6 ~ "Middle Eastern", 
+      race == 7 ~ "Two or more races", 
+      race == 8 ~ "Other"
+    ),
+    race = factor(race, levels = c(
+        "White","Black","Hispanic", "Asian", "Native American", "Middle Eastern",  "Two or more races", "Other")
+    ), 
+    presvote16post = case_when(
+      presvote16post == 1 ~ "Hillary Clinton",
+      presvote16post == 2 ~ "Donald Trump",
+      presvote16post == 3 ~ "Gary Johnson",
+      presvote16post == 4 ~ "Jill Stein",
+      presvote16post == 5 ~ "Evan McMullin",
+      presvote16post == 6 ~ "Other", 
+      presvote16post == 7 ~ "Did not vote"
+    ),
+    presvote16post = factor(presvote16post, levels = c(
+        "Hillary Clinton","Donald Trump","Gary Johnson", "Jill Stein", "Evan McMullin", "Other", "Did not vote")
+    ),
+    religpew = case_when(
+      religpew == 1 ~ "Protestant",
+      religpew == 2 ~ "Roman Catholic",
+      religpew == 3 ~ "Mormon",
+      religpew == 4 ~ "Eastern or Greek Orthodox",
+      religpew == 5 ~ "Jewish",
+      religpew == 6 ~ "Muslim", 
+      religpew == 7 ~ "Buddhist", 
+      religpew == 8 ~ "Hindu", 
+      religpew == 9 ~ "Atheist", 
+      religpew == 10 ~ "Agnostic",
+      religpew == 11 ~ "Nothing in Particular",
+      religpew == 12 ~ "Something else"
+    ),
+    religpew = factor(
+      religpew,
+      levels = c(
+        "Protestant",
+        "Roman Catholic",
+        "Mormon",
+        "Eastern or Greek Orthodox",
+        "Jewish",
+        "Muslim", 
+        "Buddhist", 
+        "Hindu", "Atheist", "Agnostic", "Nothing in Particular", "Something else"
+      )
+    ), 
+    edloan = case_when(
+      edloan == 1 ~ "Yes",
+      edloan == 2 ~ "No",
+      edloan == 3 ~ "Other",
+      edloan == 4 ~ "Don't know",
+      edloan == 5 ~ "None of these",
+    ),
+    edloan = factor(edloan, levels = c( "Yes","No", "Other", "Don't Know", "None of these")
+    ) #########ADD input state it if you want to (sakura)
+  ) |>
+  select(votereg, employ, immstat, dualcit, gender, birthyr, race, inputstate, presvote16post, religpew, edloan)
+
 ```
@@
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+- included variables of gender, birth year, race, registered state, employment, education loans, immigration status, dual-citizenship, religion, and 2016 US vote
 
-```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
+# Model
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
+The variable 'presvote16post' captures the 2016 US Presidential votes of individuals in the United States.
+Due to the minimal votes for Gary Johnson, Jill Stein, Evan McMullin, and all other candidates any votes not attributed to 'Hillary Clinton' or 'Donald Trump' were categorized as 'Other'.
 
-Talk more about it.
+We have modeled the following logisitc regression in @fig-votes:
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
+```{r, fig.height=4, fig.width=8}
 #| echo: false
@@
 #| message: false
+#| label: fig-votes
+#| fig-cap: 'Logistic regression of 2016 US Presidential votes comparing parameters of gender and employment status'
 
-analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
+#logisitc regression for binary outcome: YES OR NO, THIS OR THAT
 
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
+#choose values "Hillary Clinton" and "Donald Trump", all other values as "Other"
+ces2020_votes <- ces2020 %>%
+  filter(presvote16post != 'Did not vote') %>%
+  mutate(presvote16post = case_when(
+    presvote16post == "Hillary Clinton" ~ "Hillary Clinton",
+    presvote16post == "Donald Trump" ~ "Donald Trump",
+    presvote16post == "Gary Johnson" ~ "Other",
+    presvote16post == "Jill Stein" ~ "Other",
+    presvote16post == "Evan McMullin" ~ "Other",
+    presvote16post == "Other" ~ "Other"
+  ))
+
+#plot 2016 votes against employment status and gender
+#SO MANY NA VALUES???
+ces2020_votes %>%
+  ggplot(aes(x= employ, fill=presvote16post)) +
+  stat_count(position = "dodge") +
+  facet_wrap(facets = vars(gender)) +
   theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
+  coord_flip() +
+  labs(x ="Employment status",
+      y = "Count of votes",
+      fill = "Vote in 2016")
+
 ```
-
-Talk way more about it.
-
-# Model
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
 
@@
 
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
 ```
@@
 ## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
 
@@
 
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
 ```

### 2024-03-13 paper/paper.qmd
graphs updated

@@
 library(tidyverse)
+library(patchwork)
 ```
@@
     presvote16post = case_when(
-      presvote16post == 1 ~ "Hillary Clinton",
-      presvote16post == 2 ~ "Donald Trump",
+      presvote16post == 1 ~ "Donald Trump",
+      presvote16post == 2 ~ "Hillary Clinton",
       presvote16post == 3 ~ "Gary Johnson",
@@
 
-You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
-
-The remainder of this paper is structured as follows. @sec-data....
-
 # Data {#sec-data}
@@
   facet_wrap(facets = vars(gender)) +
+  theme_minimal() +
+  coord_flip() +
+  labs(x ="Employment status",
+      y = "Count of votes",
+      fill = "Vote in 2016")
+
+```
+
+```{r, fig.height=4, fig.width=8}
+#| echo: false
+#| warning: false
+#| message: false
+
+ces2020_votes <- ces2020_votes %>%
+  filter(!is.na(dualcit))
+
+ces2020_votes %>% 
+  ggplot(aes(x= immstat, fill=presvote16post)) +
+  stat_count(position = "dodge") +
+  facet_wrap(facets = vars(dualcit), scales = "free_x") +
+  theme_minimal() +
+  coord_flip() +
+  labs(x ="Employment status",
+      y = "Count of votes",
+      fill = "Vote in 2016")
+
+```
+
+```{r, fig.height=4, fig.width=8}
+#| echo: false
+#| warning: false
+#| message: false
+
+ces2020_votes %>% 
+  ggplot(aes(x= race, fill=presvote16post)) +
+  stat_count(position = "dodge") +
+  facet_wrap(facets = vars(gender), scales = "free_x") +
   theme_minimal() +

### 2024-03-14 paper/paper.qmd
Made a linear model

@@
 library(patchwork)
+library(gridExtra)
 ```
@@
       fill = "Vote in 2016")
+```
 
+```{r, fig.height=4, fig.width=8}
 ```
@@
 
+```{r}
+# Calculate percentages
+percentage_data <- ces2020_votes %>%
+  group_by(birthyr) %>%
+  summarise(HillaryClinton = sum(presvote16post == "Hillary Clinton") / n() * 100,
+            DonaldTrump = sum(presvote16post == "Donald Trump") / n() * 100,
+            Other = sum(presvote16post == "Other") / n() * 100,
+            DidNotVote = sum(presvote16post == "Did not Vote") / n() * 100) %>%
+  ungroup()
+
+# Reshape data to long format
+percentage_data_long <- pivot_longer(percentage_data, cols = -birthyr, names_to = "Candidate", values_to = "Percentage")
+
+plot_hillary <- ggplot(percentage_data_long %>% filter(Candidate == "HillaryClinton"), aes(x = birthyr, y = Percentage)) +
+  geom_point() + 
+  geom_smooth(method = "lm", se = FALSE, color = "blue") +
+  labs(title = "Percentage for Hillary Clinton", y = "Percentage (%)") +
+  theme_minimal()
+plot_hillary
+
+plot_trump <- ggplot(percentage_data_long %>% filter(Candidate == "DonaldTrump"), aes(x = birthyr, y = Percentage)) +
+  geom_point() + 
+  geom_smooth(method = "lm", se = FALSE, color = "blue") +
+  labs(title = "Percentage for Donald Trump", y = "Percentage (%)") +
+  theme_minimal()
+plot_trump
+
+plot_other <- ggplot(percentage_data_long %>% filter(Candidate == "Other"), aes(x = birthyr, y = Percentage)) +
+  geom_point() + 
+  geom_smooth(method = "lm", se = FALSE, color = "blue") +
+  labs(title = "Percentage for Other", y = "Percentage (%)") +
+  theme_minimal()
+plot_other
+
+```
+
 ### Model justification
@@
 
-# Appendix {-}
+# Appendix {.unnumbered}
 

### 2024-03-14 paper/paper.qmd
just updated the references and finished the data section but let me know if there are more variables we need to add in the explanation

--- 2024-03-13 paper/paper.qmd
@@
 
+## Data Used
+This paper was modeled with the help of R [@citeR] along with other useful packages like tidyverse [@tidy] (which includes graphing functions like ggplot2), patchwork [@patch].
+There are parts of the code which were guided by Rohan Alexander's Telling Stories with fire [@rohan] chapter 13 section 13.2.2.
+
+## Variables inspected
+Starting off we look at the column votereg and CS20_410.
+They represent the number of people that registered to vote and who they voted for, respectively.
+The columns where the votereg was 2, meant that the person was not registered to vote so we essentially cleaned up those values to only look at people who did vote.
+The next variable we isolated was 2016pres which tells us, who Americans voted for during the 2016 United States Presidential Election.
+This is an important variable as we will be able to see if the citizens of america were satisfied with the service that the previous government provided.
+Then we look at gender as well as employment.
+Both gender and employment shows us if there is a correlation between a certain parties views vs the demographic they are a part of and represent.
+Gender contains 2 values (male and female) while, employment has 9 values; full time worker, part time worker, laid off, unemployed, retired, permanently disabled individual, Homemaker, Student or Other.
+
+## Data Explained
+There could have been many other similar datasets that could have been used for this project for example we could have chosen to look at the census and election data for Canada.
+However, our group decided that because part of the analysis was done in @rohan, there were still many other variables that we could explore as we dive furthur into the 2020 presidential election and try to intrepret if there are any correlations between the variables and the result.
+Our team found it interesting to see all the variables that were collected by the US government and the correlations we saw during the analysis process; where there most definitely was a positive correlation between each variable and the outcome of the votes.
+Although we are analyzing the 2020 election that has already taken place, the analysis we do in the later sections are believed to apply to the 2024 elections happening this year.
+This is enough reason for us and the reader to dive into the patterns that exist with this large dataset.
+
 # Model

### 2024-03-14 paper/paper.qmd
commiting the small changes before i pull

@@
 
-## Data Explained
+## The Destination to Reach with the Data
 There could have been many other similar datasets that could have been used for this project for example we could have chosen to look at the census and election data for Canada.

### 2024-03-15 paper/paper.qmd
model 2 anlaysis and small fixes

@@
 library(gridExtra)
+library(rstanarm)
+library(modelsummary)
 ```
@@
   ) |>
-  select(votereg, employ, immstat, dualcit, gender, birthyr, race, inputstate, presvote16post, religpew, edloan)
+  select(voted_for, votereg, employ, immstat, dualcit, gender, birthyr, race, inputstate, presvote16post, religpew, edloan)
 
@@
 # Introduction
+This research examines the voting patterns in the 2016 and 2020 US Presidential elections.
+We will be looking at data collected by the Cooperative Election Study and accessed through the Harvard University Database **cite**.
+The anaylsis is based on a representative sample of 61,000 American adults, which provides detailed information about each individuals gender, birth year, race, registered state, employment, education loans, immigration status, dual-citizenship, religion, and 2016 and 2020 Presidential vote.
+The goal of this study is to use relevant variables from the electoral data to investigate patterns, trend and predictions regrading American electoral preferences from 2016 and 2020.
 
 # Data {#sec-data}
-
--   included variables of gender, birth year, race, registered state, employment, education loans, immigration status, dual-citizenship, religion, and 2016 US vote
 
@@
 ## Variables inspected
-Starting off we look at the column votereg and CS20_410.
-They represent the number of people that registered to vote and who they voted for, respectively.
-The columns where the votereg was 2, meant that the person was not registered to vote so we essentially cleaned up those values to only look at people who did vote.
-The next variable we isolated was 2016pres which tells us, who Americans voted for during the 2016 United States Presidential Election.
-This is an important variable as we will be able to see if the citizens of america were satisfied with the service that the previous government provided.
-Then we look at gender as well as employment.
-Both gender and employment shows us if there is a correlation between a certain parties views vs the demographic they are a part of and represent.
-Gender contains 2 values (male and female) while, employment has 9 values; full time worker, part time worker, laid off, unemployed, retired, permanently disabled individual, Homemaker, Student or Other.
+Starting off, we examine the columns 'votereg' and 'voted_for'.
+They represent the number of persons that registered to vote and which candidate they voted for in 2020, respectively.
+We filtered out the rows with a 'votereg' value of 2, which indicated unregistered voters, to focus exclusively on individuals who were registered to vote.
+We then focused on the 'presvote16post' variable, which reveals the candidates Americans voted for in the 2016 United States Presidential Election.
+This is an important variable as it enables us to assess whether American citizens were satisfied with the service that the previous government provided.
+Next we look at 'gender' as well as 'employment'.
+Both 'gender' and 'employment' shows us if there is a correlation between certain parties views versus the demographic they represent.
+'Gender' contains 2 values (male and female) while, employment has 9 values; full time worker, part time worker, laid off, unemployed, retired, permanently disabled individual, Homemaker, Student or Other.
+We also explore the variable 'immstat' which represents the immigration status of the of individual represented by one of the following: immigrant and citizen, immigrant not citizen, born in US, but parent(s) immigrant, parent and I born in US but grandparent(s) immigrant, or all born in US.
 
 ## The Destination to Reach with the Data
-There could have been many other similar datasets that could have been used for this project for example we could have chosen to look at the census and election data for Canada.
-However, our group decided that because part of the analysis was done in @rohan, there were still many other variables that we could explore as we dive furthur into the 2020 presidential election and try to intrepret if there are any correlations between the variables and the result.
+There could have been many other similar data sets that could have been used for this project for example we could have chosen to look at the census and election data for Canada.
+However, our group decided that because part of the analysis was done in @rohan, there were still many other variables that we could explore as we dive further into the 2020 presidential election and try to interpret if there are any correlations between the variables and the result.
 Our team found it interesting to see all the variables that were collected by the US government and the correlations we saw during the analysis process; where there most definitely was a positive correlation between each variable and the outcome of the votes.
 Although we are analyzing the 2020 election that has already taken place, the analysis we do in the later sections are believed to apply to the 2024 elections happening this year.
-This is enough reason for us and the reader to dive into the patterns that exist with this large dataset.
+This is enough reason for us and the reader to dive into the patterns that exist with this large data set.
 
-# Model
+# Model 1
+Initially, we filter the data to remove to exclude individuals who did not vote for Hillary Clinton or Donald Trump, due to the minimal votes for Gary Johnson, Jill Stein, Evan McMullin, and all other candidates which are insignificant to this paper.
+We first analyze the birth year of voters.
 
-The variable 'presvote16post' captures the 2016 US Presidential votes of individuals in the United States.
-Due to the minimal votes for Gary Johnson, Jill Stein, Evan McMullin, and all other candidates any votes not attributed to 'Hillary Clinton' or 'Donald Trump' were categorized as 'Other'.
-
-We have modeled the following logisitc regression in @fig-votes:
-
-```{r, fig.height=4, fig.width=8}
-#| echo: false
-#| warning: false
-#| message: false
-#| label: fig-votes
-#| fig-cap: 'Logistic regression of 2016 US Presidential votes comparing parameters of gender and employment status'
-
-#logisitc regression for binary outcome: YES OR NO, THIS OR THAT
-
-#choose values "Hillary Clinton" and "Donald Trump", all other values as "Other"
-ces2020_votes <- ces2020 %>%
-  filter(presvote16post != 'Did not vote') %>%
-  mutate(presvote16post = case_when(
-    presvote16post == "Hillary Clinton" ~ "Hillary Clinton",
-    presvote16post == "Donald Trump" ~ "Donald Trump",
-    presvote16post == "Gary Johnson" ~ "Other",
-    presvote16post == "Jill Stein" ~ "Other",
-    presvote16post == "Evan McMullin" ~ "Other",
-    presvote16post == "Other" ~ "Other"
-  ))
-
-#plot 2016 votes against employment status and gender
-#SO MANY NA VALUES???
-ces2020_votes %>%
-  ggplot(aes(x= employ, fill=presvote16post)) +
-  stat_count(position = "dodge") +
-  facet_wrap(facets = vars(gender)) +
-  theme_minimal() +
-  coord_flip() +
-  labs(x ="Employment status",
-      y = "Count of votes",
-      fill = "Vote in 2016")
-```
-
-```{r, fig.height=4, fig.width=8}
-```
-
-```{r, fig.height=4, fig.width=8}
+```{r}
 #| echo: false
@@
 
-ces2020_votes <- ces2020_votes %>%
-  filter(!is.na(dualcit))
+# only choose values "Hillary Clinton" and "Donald Trump"
+ces2020_votes <- ces2020 %>%
+  filter(presvote16post != 'Did not vote') %>%
+  filter(presvote16post %in% c("Hillary Clinton", "Donald Trump"))
 
-ces2020_votes %>% 
-  ggplot(aes(x= immstat, fill=presvote16post)) +
-  stat_count(position = "dodge") +
-  facet_wrap(facets = vars(dualcit), scales = "free_x") +
-  theme_minimal() +
-  coord_flip() +
-  labs(x ="Employment status",
-      y = "Count of votes",
-      fill = "Vote in 2016")
-
-```
-
-```{r, fig.height=4, fig.width=8}
-#| echo: false
-#| warning: false
-#| message: false
-
-ces2020_votes %>% 
-  ggplot(aes(x= race, fill=presvote16post)) +
-  stat_count(position = "dodge") +
-  facet_wrap(facets = vars(gender), scales = "free_x") +
-  theme_minimal() +
-  coord_flip() +
-  labs(x ="Employment status",
-      y = "Count of votes",
-      fill = "Vote in 2016")
-
-```
-
-```{r}
 # Calculate percentages
@@
 
+##porbably dont need this - cris
 plot_other <- ggplot(percentage_data_long %>% filter(Candidate == "Other"), aes(x = birthyr, y = Percentage)) +
@@
 plot_other
+
+```
+In this model, we conduct an examination of the relationship between voters' birth year and their gender for the 2016 and 2020 Presidential elections.
+This analysis is visualized by plotting a histogram that separate female voters on the left and male voters on the right, with voters' birth years measured along the x-axis, which ranges from 1925 to 2002.
+The y-axis quantifies the voter turnout for the year.
+For clarity ad symbolic representation, the colour blue was chosen to represent the Democratic candidates --Hillary Clinton for the 2016 election, and Joe Biden for the 2020 election, while red was chosen to the Represent the Republican candidate, Donald Trump, who sought the presidency in both terms.
+@fig-genderyr2016 present the distribution of votes in 2016, and @fig-genderyr2020 presents the data from the 2020 election.
+
+We have modeled the following logistic regression in the graphs:
+
+$$
+\begin{aligned} 
+y_i|\pi_i  &\sim \mbox{Bern}(\pi_i) \\
+\mbox{logit}(\pi_i) &= \beta_0 + \beta_1 \times \mbox{gender}_i + \beta_2\times \mbox{immigration status}_i\\
+\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_2 &\sim \mbox{Normal}(0, 2.5)
+\end{aligned}
+$$
+
+```{r, fig.height=6, fig.width=11}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-genderyr2016
+#| fig-cap: 'Logistic regression of 2016 US Presidential votes comparing parameters of gender and immigration status'
+
+my_palette <- c("Hillary Clinton" = "skyblue3", "Donald Trump" = "indianred3")
+my_palette2 <- c("Biden" = "skyblue3", "Trump" = "indianred3")
+
+ces2020_votes %>% 
+  ggplot(aes(x= birthyr, fill=presvote16post)) +
+  geom_histogram(binwidth = 2, position = "dodge", alpha = 0.8) + 
+  facet_wrap(facets = vars(gender), scales = "free_x") +
+  theme_minimal() +
+  scale_fill_manual(values = my_palette) + 
+  guides(fill = guide_legend(reverse = TRUE)) +
+  labs(x ="Birth Year",
+      y = "Count of votes",
+      fill = "Vote in 2016")
+```
+
+```{r, fig.height=6, fig.width=11}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-genderyr2020
+#| fig-cap: 'Logistic regression of 2020 US Presidential votes comparing parameters of gender and immigration status'
+
+ces2020_votes %>% 
+  ggplot(aes(x= birthyr, fill=voted_for)) +
+  geom_histogram(binwidth = 2, position = "dodge", alpha = 0.8) + 
+  facet_wrap(facets = vars(gender), scales = "free_x") +
+  theme_minimal() +
+  scale_fill_manual(values = my_palette2) + 
+  guides(fill = guide_legend(reverse = TRUE)) +
+  labs(x ="Birth Year",
+      y = "Count of votes",
+      fill = "Vote in 2020")
+
+```
+In this model, we conduct an examination of the relationship between voters' birth year and their gender for the 2016 and 2020 Presidential elections.
+This analysis is visualized by plotting a histogram that separate female voters on the left and male voters on the right, with voters' birth years measured along the x-axis, which ranges from 1925 to 2002.
+The y-axis quantifies the voter turnout for the year.
+For clarity ad symbolic representation, the colour blue was chosen to represent the Democratic candidates --Hillary Clinton for the 2016 election, and Joe Biden for the 2020 election, while red was chosen to the Represent the Republican candidate, Donald Trump, who sought the presidency in both terms.
+@fig-genderyr2016 present the distribution of votes in 2016, and @ifg-genderyr2020 presents the data from the 2020 election.
+
+# Model 2
+
+```{r, fig.align='right', fig.height=5, fig.width=11}
+#| echo: false
+#| warning: false
+#| message: false
+
+#plot for 2016 votes
+ces2020_votes %>% 
+  ggplot(aes(x= immstat, fill=presvote16post)) +
+  stat_count(position = "dodge") +
+  facet_wrap(facets = vars(gender), scales = "free_x") +
+  theme_minimal() +
+  coord_flip() +
+  scale_fill_manual(values = my_palette) +
+  guides(fill = guide_legend(reverse = TRUE)) +
+  labs(x ="Immigration status",
+      y = "Count of votes",
+      fill = "Vote in 2016")
+
+#plot for 2020 votes
+ces2020_votes %>% 
+  ggplot(aes(x= immstat, fill=voted_for)) +
+  stat_count(position = "dodge") +
+  facet_wrap(facets = vars(gender), scales = "free_x") +
+  theme_minimal() +
+  coord_flip() +
+  scale_fill_manual(values = my_palette2) + 
+  guides(fill = guide_legend(reverse = TRUE)) +
+  labs(x ="Immigration status",
+      y = "Count of votes",
+      fill = "Vote in 2020")
+
+#trump republican hillary and bidne were democrat 
+#male born in the US are a 50/50, everyone else changed minds
+```
+```{r}
 

### 2024-03-16 paper/paper.qmd
more. analysis model. 2

@@
 library(modelsummary)
+library(dplyr)
+library(lubridate)
+library(kableExtra)
+library(knitr)
 ```
@@
 
-# Model 1
+# Models
+
+## Model 1
 Initially, we filter the data to remove to exclude individuals who did not vote for Hillary Clinton or Donald Trump, due to the minimal votes for Gary Johnson, Jill Stein, Evan McMullin, and all other candidates which are insignificant to this paper.
@@
 ```
+## Model 2
 In this model, we conduct an examination of the relationship between voters' birth year and their gender for the 2016 and 2020 Presidential elections.
@@
 @fig-genderyr2016 present the distribution of votes in 2016, and @fig-genderyr2020 presents the data from the 2020 election.
+
+```{r, fig.height=6, fig.width=11}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-genderyr2016
+#| fig-cap: "Logistic regression of 2016 US Presidential votes comparing parameters of gender and immigration status"
+
+my_palette <- c("Hillary Clinton" = "skyblue3", "Donald Trump" = "indianred3")
+my_palette2 <- c("Biden" = "skyblue3", "Trump" = "indianred3")
+
+ces2020_votes %>% 
+  ggplot(aes(x= birthyr, fill=presvote16post)) +
+  geom_histogram(binwidth = 2, position = "dodge", alpha = 0.8) + 
+  facet_wrap(facets = vars(gender), scales = "free_x") +
+  theme_minimal() +
+  scale_fill_manual(values = my_palette) + 
+  guides(fill = guide_legend(reverse = TRUE)) +
+  labs(x ="Birth year",
+      y = "Count of votes",
+      fill = "Vote in 2016")
+```
+We notice in @fig-genderyr2016 the graph displays two distinct high points for the number of female Republican voters, with the peak around 1960 being the most pronounced, followed by another around 1980.
+There are similar peaks in the graph for female Democratic voters, but with the overall count being considerably lower.
+We also notice that the Democratic party received slightly more votes from the older demographic, where as the younger demographic greatly preferred the Republican party.
+In comparison, the graph displaying the male votes has a more balanced distribution, and similar to the female voters, the older and younger demographic preferred the Democratic Party and the Republican Party, respectively.
+It is worthy to mention that there is a higher count of women than men.
+
+```{r, fig.height=6, fig.width=11}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-genderyr2020
+#| fig-cap: 'Logistic regression of 2020 US Presidential votes comparing parameters of gender and immigration status'
+
+#plot for 2020 votes by 
+ces2020_votes %>% 
+  ggplot(aes(x= birthyr, fill=voted_for)) +
+  geom_histogram(binwidth = 2, position = "dodge", alpha = 0.8) + 
+  facet_wrap(facets = vars(gender), scales = "free_x") +
+  theme_minimal() +
+  scale_fill_manual(values = my_palette2) + 
+  guides(fill = guide_legend(reverse = TRUE)) +
+  labs(x ="Birth Year",
+      y = "Count of votes",
+      fill = "Vote in 2020")
+
+```
+We notice in @fig-genderyr2020 that the graphs bear a strong resemblance to the distributions of the 2016 votes depicted in @fig-genderyr2016, but with the parties reversed on the graph.
+We are interested in the ratios of these graphs.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: tbl-ratio2016
+
+ces2020_votes_2016 <- ces2020_votes %>%
+  mutate(birth_decade = (birthyr %/% 10) * 10,
+    trump_vote = ifelse(presvote16post == "Donald Trump", 1, 0),
+    biden_vote = ifelse(presvote16post == "Hillary Clinton", 1, 0))
+
+ces2020_votes_2016 <- ces2020_votes_2016 %>%
+  filter(!(birth_decade %in% c(1920, 2000)))
+
+votes_count <- ces2020_votes_2016 %>%
+  group_by(birth_decade, gender) %>%
+  summarise(biden_vote_total = sum(biden_vote), trump_vote_total = sum(trump_vote), .groups = "drop")
+
+votes_count <- votes_count %>%
+  mutate(vote_ratio = biden_vote_total / trump_vote_total)
+
+votes_count_female <- votes_count %>% 
+  filter(gender == "Female")
+votes_count_male <- votes_count %>% 
+  filter(gender == "Male")
+
+votes_count_female <- select(votes_count_female, -gender)
+votes_count_male <- select(votes_count_male, -gender)
+
+colnames(votes_count_female) <- c("Birth Decade", "Total Clinton Votes", "Total Trump Votes", "Ratio")
+colnames(votes_count_male) <- c("Birth Decade", "Total Clinton Votes", "Total Trump Votes", "Ratio")
+
+votes_2020_f <- kable(votes_count_female, 
+      caption = "Ratio of 2016 Votes for Clinton to Trump by Birth Decade (Females)",
+      digits = 2)
+
+votes_2020_m <- kable(votes_count_male, 
+      caption = "Ratio of 2016 Votes for Clinton to Trump by Birth Decade (Males)",
+      digits = 2)
+
+kables(list(votes_2020_f, votes_2020_m))
+#If the ratio is greater than 1, it means that within the group, Biden has more votes than Trump
+#if the ratio is less than 1, Trump has more votes within the group
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: tbl-ratio2020
+
+ces2020_votes_2020 <- ces2020_votes %>%
+  mutate(birth_decade = (birthyr %/% 10) * 10,
+    trump_vote = ifelse(voted_for == "Trump", 1, 0),
+    biden_vote = ifelse(voted_for == "Biden", 1, 0))
+
+ces2020_votes_2020 <- ces2020_votes_2020 %>%
+  filter(!(birth_decade %in% c(1920, 2000)))
+
+votes_count <- ces2020_votes_2020 %>%
+  group_by(birth_decade, gender) %>%
+  summarise(biden_vote_total = sum(biden_vote), trump_vote_total = sum(trump_vote), .groups = "drop")
+
+votes_count <- votes_count %>%
+  mutate(vote_ratio = biden_vote_total / trump_vote_total)
+
+votes_count_female <- votes_count %>% 
+  filter(gender == "Female")
+votes_count_male <- votes_count %>% 
+  filter(gender == "Male")
+
+votes_count_female <- select(votes_count_female, -gender)
+votes_count_male <- select(votes_count_male, -gender)
+
+colnames(votes_count_female) <- c("Birth Decade", "Total Biden Votes", "Total Trump Votes", "Ratio")
+colnames(votes_count_male) <- c("Birth Decade", "Total Biden Votes", "Total Trump Votes", "Ratio")
+
+votes_2020_f <- kable(votes_count_female, 
+      caption = "Ratio of 2020 Votes for Biden to Trump by Birth Decade (Females)",
+      digits = 2)
+
+votes_2020_m <- kable(votes_count_male, 
+      caption = "Ratio of 2020 Votes for Biden to Trump by Birth Decade (Males)",
+      digits = 2)
+
+kables(list(votes_2020_f, votes_2020_m))
+#If the ratio is greater than 1, it means that within the group, Biden has more votes than Trump
+#if the ratio is less than 1, Trump has more votes within the group
+```
+
+We analyze the ratio of Democratic to Republic votes in 2016 in @tbl-ratio2016 and in 2020 in @tbl-ratio2020.
+We divide respondents by birth decade, and we list the total number of votes for the Democratic Party and thee Republican Party, then list the ratio of the two.
+The decades 1920 and 2000 were omitted since the number of respondents in those Birth Decades were very low.
+The value of the ratio in centered at 1, where if the values is greater than 1, then within that group, the Democratic Party has more votes than the Republican Party.
+Where the values are less than 1, the Republican Party has more votes.
+When the ratio is around 1, the number of votes is approximately equal.
+
+We notice that in 2016, the ratio of votes decreases as birth decade increases for both women and men voters.
+The opposite happens in 2020, where the ratio of votes increases as birth decade increases.
+
+## Model 3
 
@@
 $$
-
-```{r, fig.height=6, fig.width=11}
-#| echo: false
-#| warning: false
-#| message: false
-#| label: fig-genderyr2016
-#| fig-cap: 'Logistic regression of 2016 US Presidential votes comparing parameters of gender and immigration status'
-
-my_palette <- c("Hillary Clinton" = "skyblue3", "Donald Trump" = "indianred3")
-my_palette2 <- c("Biden" = "skyblue3", "Trump" = "indianred3")
-
-ces2020_votes %>% 
-  ggplot(aes(x= birthyr, fill=presvote16post)) +
-  geom_histogram(binwidth = 2, position = "dodge", alpha = 0.8) + 
-  facet_wrap(facets = vars(gender), scales = "free_x") +
-  theme_minimal() +
-  scale_fill_manual(values = my_palette) + 
-  guides(fill = guide_legend(reverse = TRUE)) +
-  labs(x ="Birth Year",
-      y = "Count of votes",
-      fill = "Vote in 2016")
-```
-
-```{r, fig.height=6, fig.width=11}
-#| echo: false
-#| warning: false
-#| message: false
-#| label: fig-genderyr2020
-#| fig-cap: 'Logistic regression of 2020 US Presidential votes comparing parameters of gender and immigration status'
-
-ces2020_votes %>% 
-  ggplot(aes(x= birthyr, fill=voted_for)) +
-  geom_histogram(binwidth = 2, position = "dodge", alpha = 0.8) + 
-  facet_wrap(facets = vars(gender), scales = "free_x") +
-  theme_minimal() +
-  scale_fill_manual(values = my_palette2) + 
-  guides(fill = guide_legend(reverse = TRUE)) +
-  labs(x ="Birth Year",
-      y = "Count of votes",
-      fill = "Vote in 2020")
-
-```
-In this model, we conduct an examination of the relationship between voters' birth year and their gender for the 2016 and 2020 Presidential elections.
-This analysis is visualized by plotting a histogram that separate female voters on the left and male voters on the right, with voters' birth years measured along the x-axis, which ranges from 1925 to 2002.
-The y-axis quantifies the voter turnout for the year.
-For clarity ad symbolic representation, the colour blue was chosen to represent the Democratic candidates --Hillary Clinton for the 2016 election, and Joe Biden for the 2020 election, while red was chosen to the Represent the Republican candidate, Donald Trump, who sought the presidency in both terms.
-@fig-genderyr2016 present the distribution of votes in 2016, and @ifg-genderyr2020 presents the data from the 2020 election.
-
-# Model 2
 
@@
 ```
+
 ```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
+set.seed(302)
+
+ces2020_reduced <- 
+  ces2020_votes |> 
+  slice_sample(n = 1000)
+
+political_preferences <-
+  stan_glm(
+    voted_for ~ gender + immstat,
+    data = ces2020_reduced,
+    family = binomial(link = "logit"),
+    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+    prior_intercept = 
+      normal(location = 0, scale = 2.5, autoscale = TRUE),
+    seed = 302
+  )
+
+saveRDS(
+  political_preferences,
+  file = "political_preferences.rds"
+)
+
+political_preferences <-
+  readRDS(file = "political_preferences.rds")
+
+modelsummary(
+  list(
+    "Support Biden" = political_preferences
+  ),
+  statistic = "mad"
+  )
 
 ```
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
 

### 2024-03-16 paper/paper.qmd
commiting with cleaned data (reduced var) and intro and abstract

@@
 abstract: |
- Some abstact
+ This data uses the US 2020 Census results that is in the Harvard Database in order to analyse if there is a correltaion between the election numbers in 2020 and other factors.
+The factors include immigration status, gender, birth year, and results from the 2016 election.
+It was found that there does exist a correlation between the outcomes of the 2020 election and the above variables.
+These are important results as the US is a country that si considered a hub, due to all the external relations they have with multiple countries.
+These results will also help predict who may win the 2024 elections.
 include-in-header: 
@@
   ) |>
-  select(votereg, CC20_410, employ, immstat, dualcit, gender, birthyr, race, inputstate, presvote16post, religpew, edloan)
+  select(votereg, CC20_410, immstat, gender, birthyr, presvote16post)
 
@@
         "CC20_410" = col_integer(),
-        "employ" = col_integer(),
-        "dualcit" = col_integer(), 
+        "immstat" = col_integer(), 
         "gender" = col_integer(), 
         "birthyr" = col_integer(),  
-        "race" = col_integer(), 
-        "inputstate" = col_integer(), 
         "presvote16post" = col_integer(), 
-        "religpew" = col_integer(), 
-        "edloan" = col_integer()
       )
@@
     gender = if_else(gender == 1, "Male", "Female"),
-    employ = case_when(
-      employ == 1 ~ "Full-time",
-      employ == 2 ~ "Part-time",
-      employ == 3 ~ "Laid off",
-      employ == 4 ~ "Unemployed",
-      employ == 5 ~ "Retired",
-      employ == 6 ~ "Permannetly disabled", 
-      employ == 7 ~ "Homemaker", 
-      employ == 8 ~ "Student", 
-      employ == 9 ~ "Other"
-    ),
-    employ = factor(
-      employ,
-      levels = c(
-        "Full-time",
-        "Part-time",
-        "laid off",
-        "unemployed",
-        "retired",
-        "permannetly disabled", 
-        "Homemaker", 
-        "Student", "Other"
-      )
-    ), 
     immstat = case_when(
@@
     immstat = factor(immstat, levels = c( "immigrant and citizen","immigrant not citizen", "born in US, but parent(s) immigrant",     "parent and I born in US but grandparent(s) immigrant", "all born in US")
-    ),
-    dualcit = if_else(dualcit == 1, "Yes", "No"),
-    race = case_when(
-      race == 1 ~ "White",
-      race == 2 ~ "Black",
-      race == 3 ~ "Hispanic",
-      race == 4 ~ "Asian",
-      race == 5 ~ "Native American",
-      race == 6 ~ "Middle Eastern", 
-      race == 7 ~ "Two or more races", 
-      race == 8 ~ "Other"
-    ),
-    race = factor(race, levels = c(
-        "White","Black","Hispanic", "Asian", "Native American", "Middle Eastern",  "Two or more races", "Other")
     ),
@@
         "Hillary Clinton","Donald Trump","Gary Johnson", "Jill Stein", "Evan McMullin", "Other", "Did not vote")
-    ),
-    religpew = case_when(
-      religpew == 1 ~ "Protestant",
-      religpew == 2 ~ "Roman Catholic",
-      religpew == 3 ~ "Mormon",
-      religpew == 4 ~ "Eastern or Greek Orthodox",
-      religpew == 5 ~ "Jewish",
-      religpew == 6 ~ "Muslim", 
-      religpew == 7 ~ "Buddhist", 
-      religpew == 8 ~ "Hindu", 
-      religpew == 9 ~ "Atheist", 
-      religpew == 10 ~ "Agnostic",
-      religpew == 11 ~ "Nothing in Particular",
-      religpew == 12 ~ "Something else"
-    ),
-    religpew = factor(
-      religpew,
-      levels = c(
-        "Protestant",
-        "Roman Catholic",
-        "Mormon",
-        "Eastern or Greek Orthodox",
-        "Jewish",
-        "Muslim", 
-        "Buddhist", 
-        "Hindu", "Atheist", "Agnostic", "Nothing in Particular", "Something else"
     )
-    ), 
-    edloan = case_when(
-      edloan == 1 ~ "Yes",
-      edloan == 2 ~ "No",
-      edloan == 3 ~ "Other",
-      edloan == 4 ~ "Don't know",
-      edloan == 5 ~ "None of these",
-    ),
-    edloan = factor(edloan, levels = c( "Yes","No", "Other", "Don't Know", "None of these")
-    ) #########ADD input state it if you want to (sakura)
   ) |>
-  select(voted_for, votereg, employ, immstat, dualcit, gender, birthyr, race, inputstate, presvote16post, religpew, edloan)
+  select(voted_for, votereg, immstat, gender, birthyr, presvote16post)
 
@@
 # Introduction
+
+The United States is one of the leading countries in export, imports and almost everything economic and socially related.
+The US contributes to the worlds economy by 20% despite the fact that they contain 5% of the populations[@USTradeRep].
+This makes the US very relevant to not only national news but also international news.
+This is why the US presidential elections are broadcaster to worldwide ever election term.
+The results of the election not only affect American citizens but also external affairs related to the country.
+
+The United States of America is a democratic government which means that they hold election every 4 years[@usc_presidential_election].
+The 2 parties that historically run against each other with the most votes are the Republican party and the Democratic Party.
+The republicans are often associated with conservative beliefs and values such as views opposing abortion and privatization to save their economy.
+The Democrats, on the other hand, are often associated with liberal views such as social welfare programs and higher taxes to support the government aids provided to citizens[@britannica_dem_vs_rep].
+There are many other parties such as Libertarian Party, Green Party, Constitution Party and other independent candidates but because majority of the votes goes to the 2 parties the others are often over looked.
+The legal voting age is 18 in the states and you must be a registered voter in order to take part of the election which means non-citizens are not taken into account.
+Students are not as well.
+Also there seems to be a 66% voter turnout which means the remaining 34% decided not to vote [@DeSilver_2022].
+The 2024 elections is also coming up this year meaning that analyzing this results may help us predict what the outcome of the election might be.
+
 This research examines the voting patterns in the 2016 and 2020 US Presidential elections.
-We will be looking at data collected by the Cooperative Election Study and accessed through the Harvard University Database **cite**.
-The anaylsis is based on a representative sample of 61,000 American adults, which provides detailed information about each individuals gender, birth year, race, registered state, employment, education loans, immigration status, dual-citizenship, religion, and 2016 and 2020 Presidential vote.
+We will be looking at data collected by the Cooperative Election Study and accessed through the Harvard University Database [@dataverse].
+The analysis is based on a representative sample of 61,000 American adults, which provides detailed information about each individuals gender, birth year, race, registered state, employment, education loans, immigration status, dual-citizenship, religion, and 2016 and 2020 Presidential vote.
 The goal of this study is to use relevant variables from the electoral data to investigate patterns, trend and predictions regrading American electoral preferences from 2016 and 2020.
@@
 There are parts of the code which were guided by Rohan Alexander's Telling Stories with fire [@rohan] chapter 13 section 13.2.2.
+The data was used from the Harvard database [@dataverse]
 
@@
 
+```
+
+```{r}
 modelsummary(
@@
   )
-
 ```

### 2024-03-16 paper/paper.qmd
agains saved scripts and refrences

@@
 #| message: false
+#| include: false
 
@@
   political_preferences,
-  file = "political_preferences.rds"
+  file = "../models/political_preferences.rds"
 )
@@
 political_preferences <-
-  readRDS(file = "political_preferences.rds")
+  readRDS(file = "../models/political_preferences.rds")
 

### 2024-03-16 scripts/00-simulate_data.R

@@ -1,11 +1,8 @@
 #### Preamble ####
-# Purpose: Simulates... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Purpose: stimulate data
+# Author: Sakura Noskor, Yan Mezhiborsky, Cristina Burca
+# Date: 13 March 2023
+# Contact: cristina.burca@mail.utoronto.ca, sakura.noskor@mail.utoronto.ca,  yan.mezhiborsky@mail.utoronto.ca
 
 
 #### Workspace setup ####
@@ -14,6 +11,33 @@ library(tidyverse)
 
 #### Simulate data ####
 # [...ADD CODE HERE...]
+
+
+# Load necessary libraries
+library(dplyr)
+
+# Generate random birth years (between 1950 and 2002)
+birth_years <- sample(1950:2002, 1000, replace = TRUE)
+
+# Simulate presidential election of 2016
+election_2016 <- sample(c("Clinton", "Trump", "Gary Johnson", "other"), 1000, replace = TRUE)
+
+# Simulate gender (50% male, 50% female)
+gender <- sample(c("Male", "Female"), 1000, replace = TRUE)
+
+# Simulate immigration status (5 categories)
+immigration_status <- sample(c("immigrant and citizen", "immigrant not citizen", "born in US, but parent(s) immigrant", "parent and I born in US but grandparent(s) immigrant", "all born in US"), 1000, replace = TRUE)
+
+# Create the synthetic dataset
+election_data <- data.frame(
+  Birth_Year = birth_years,
+  Election_2016 = election_2016,
+  Gender = gender,
+  Immigration_Status = immigration_status
+)
+
+# Print the first few rows of the dataset
+head(election_data)
 
 
 

### 2024-03-16 scripts/01-download_data.R

@@ -1,26 +1,23 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Purpose: download the data
+# Author: Sakura Noskor, Yan Mezhiborsky, Cristina Burca
+# Date: 16 March 2023
+# Contact: cristina.burca@mail.utoronto.ca, sakura.noskor@mail.utoronto.ca,  yan.mezhiborsky@mail.utoronto.ca
 
 
 #### Workspace setup ####
-library(opendatatoronto)
+library(dataverse)
 library(tidyverse)
-# [...UPDATE THIS...]
 
-#### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
+ces2020 <-
+  get_dataframe_by_name(
+    filename = "CES20_Common_OUTPUT_vv.csv",
+    dataset = "10.7910/DVN/E9N6PH",
+    server = "dataverse.harvard.edu",
+    .f = read_csv
+  ) |>
+  select(votereg, CC20_410, immstat, gender, birthyr, presvote16post)
 
-
-
-#### Save data ####
-# [...UPDATE THIS...]
-# change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(the_raw_data, "inputs/data/raw_data.csv") 
+write_csv(ces2020, "data/raw_data/ces2020.csv")
 
          

### 2024-03-16 scripts/02-data_cleaning.R

@@ -1,44 +1,58 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Purpose: clean the data
+# Author: Sakura Noskor, Yan Mezhiborsky, Cristina Burca
+# Date: 13 March 2023
+# Contact: cristina.burca@mail.utoronto.ca, sakura.noskor@mail.utoronto.ca,  yan.mezhiborsky@mail.utoronto.ca
 #### Workspace setup ####
 library(tidyverse)
 
 #### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+ces2020 <-
+  read_csv(
+    "data/raw_data/ces2020.csv",
+    col_types =
+      cols(
+        "votereg" = col_integer(),
+        "CC20_410" = col_integer(),
+        "immstat" = col_integer(), 
+        "gender" = col_integer(), 
+        "birthyr" = col_integer(),  
+        "presvote16post" = col_integer(), 
+      )
+  )
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
[…]

### 2024-03-16 scripts/04-model.R

@@ -1,11 +1,8 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Purpose: model the data
+# Author: Sakura Noskor, Yan Mezhiborsky, Cristina Burca
+# Date: 13 March 2023
+# Contact: cristina.burca@mail.utoronto.ca, sakura.noskor@mail.utoronto.ca,  yan.mezhiborsky@mail.utoronto.ca
 
 
 #### Workspace setup ####
@@ -13,25 +10,27 @@ library(tidyverse)
 library(rstanarm)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+analysis_data <- read_csv("data/analysis_data/ces2020.csv")
 
 ### Model data ####
-first_model <-
+set.seed(302)
+
+ces2020_reduced <- 
+  ces2020_votes |> 
+  slice_sample(n = 1000)
+
+political_preferences <-
   stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
+    voted_for ~ gender + immstat,
+    data = ces2020_reduced,
+    family = binomial(link = "logit"),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
+    prior_intercept = 
+      normal(location = 0, scale = 2.5, autoscale = TRUE),
+    seed = 302
   )
 
-
-#### Save model ####
 saveRDS(
-  first_model,
-  file = "models/first_model.rds"
+  political_preferences,
+  file = "models/political_preferences.rds"
 )
-
-

### 2024-03-16 paper/paper.qmd
Added Graphs for 2020 election

@@
   geom_point() + 
-  geom_smooth(method = "lm", se = FALSE, color = "blue") +
+  geom_smooth(method = "lm", se = FALSE, color = "red") +
   labs(title = "Percentage for Donald Trump", y = "Percentage (%)") +
@@
 
-##porbably dont need this - cris
-plot_other <- ggplot(percentage_data_long %>% filter(Candidate == "Other"), aes(x = birthyr, y = Percentage)) +
+```
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
+ces2020_votes <- ces2020 %>%
+  filter(voted_for != 'Did not vote') %>%
+  filter(voted_for %in% c("Biden", "Trump"))
+
+# Calculate percentages
+percentage_data_2 <- ces2020_votes %>%
+  group_by(birthyr) %>%
+  summarise(
+    `Joe Biden` = sum(voted_for == "Biden") / n() * 100,  # Enclosed in backticks
+    DonaldTrump = sum(voted_for == "Trump") / n() * 100
+  ) %>%
+  ungroup()
+
+# Reshape data to long format
+percentage_data_2_long <- pivot_longer(percentage_data_2, cols = -birthyr, names_to = "Candidate", values_to = "Percentage")
+
+# Plot for Joe Biden
+plot_biden <- ggplot(percentage_data_2_long %>% filter(Candidate == "Joe Biden"), aes(x = birthyr, y = Percentage)) +
   geom_point() + 
   geom_smooth(method = "lm", se = FALSE, color = "blue") +
-  labs(title = "Percentage for Other", y = "Percentage (%)") +
+  labs(title = "Percentage for Joe Biden", y = "Percentage (%)", x = "Birth Year") +  # Adding x-axis label
   theme_minimal()
-plot_other
+
+plot_biden
+
+# Plot for Donald Trump
+plot_trump_2 <- ggplot(percentage_data_2_long %>% filter(Candidate == "DonaldTrump"), aes(x = birthyr, y = Percentage)) +  # Match summarise naming
+  geom_point() + 
+  geom_smooth(method = "lm", se = FALSE, color = "red") +
+  labs(title = "Percentage for Donald Trump", y = "Percentage (%)", x = "Birth Year") +
+  theme_minimal()
+
+plot_trump_2
 
 ```
+
 ## Model 2

### 2024-03-16 paper/paper.qmd
.

--- 2024-03-16 paper/paper.qmd
@@
 
-## Model 1
-Initially, we filter the data to remove to exclude individuals who did not vote for Hillary Clinton or Donald Trump, due to the minimal votes for Gary Johnson, Jill Stein, Evan McMullin, and all other candidates which are insignificant to this paper.
-We first analyze the birth year of voters.
+## Aging and Its Impact on Voting
+
+In these following models we will of our research paper, we will look at the relationship between the demographic of age within the population and their voting patterns.The focus will primarily be on understanding how different age groups align themselves politically.
+By investigating the percentage of voters within each age bracket who support particular candidates, we aim to uncover nuanced insights into the demographic underpinnings of political support.Through the integration of descriptive statistics, analysis, and visual data representation we will offer a comprehensive overview on age base voting patterns.
+This will not only provide empirical evidence on the voting preferences of different ages but also help for a deeper understanding of the potential motivations behind their voting choices.
+
+### Voters in 2016
 
@@
 #| message: false
+#| label: fig-hillary2016
+#| fig-cap: "Linear Regression comparing Percentage of Votes For Hillary Clinton and Year of Birth (2016)"
 
@@
 ces2020_votes <- ces2020 %>%
-  filter(presvote16post != 'Did not vote') %>%
-  filter(presvote16post %in% c("Hillary Clinton", "Donald Trump"))
+  filter(presvote16post %in% c("Hillary Clinton", "Donald Trump", "Other", "Did not vote"))
 
@@
   group_by(birthyr) %>%
-  summarise(HillaryClinton = sum(presvote16post == "Hillary Clinton") / n() * 100,
+  summarise(
+    HillaryClinton = sum(presvote16post == "Hillary Clinton") / n() * 100,
     DonaldTrump = sum(presvote16post == "Donald Trump") / n() * 100,
     Other = sum(presvote16post == "Other") / n() * 100,
-            DidNotVote = sum(presvote16post == "Did not Vote") / n() * 100) %>%
+    DidNotVote = sum(presvote16post == "Did not vote") / n() * 100
+  ) %>%
   ungroup()
@@
 
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-trump2016
+#| fig-cap: "Linear Regression comparing Percentage of Votes For Donald Trump and Year of Birth (2016)"
+
+#Trum Plot
 plot_trump <- ggplot(percentage_data_long %>% filter(Candidate == "DonaldTrump"), aes(x = birthyr, y = Percentage)) +
   geom_point() + 
-  geom_smooth(method = "lm", se = FALSE, color = "blue") +
+  geom_smooth(method = "lm", se = FALSE, color = "red") +
   labs(title = "Percentage for Donald Trump", y = "Percentage (%)") +
@@
 plot_trump
+```
 
-##porbably dont need this - cris
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-other2016
+#| fig-cap: "Linear Regression comparing Percentage of Votes For Other Caniddates and Year of Birth (2016)"
+
 plot_other <- ggplot(percentage_data_long %>% filter(Candidate == "Other"), aes(x = birthyr, y = Percentage)) +
   geom_point() + 
-  geom_smooth(method = "lm", se = FALSE, color = "blue") +
-  labs(title = "Percentage for Other", y = "Percentage (%)") +
+  geom_smooth(method = "lm", se = FALSE, color = "green") +
+  labs(title = "Percentage for Other", x = "Birth Year", y = "Percentage (%)") +
   theme_minimal()
 plot_other
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-dnv2016
+#| fig-cap: "Linear Regression comparing Percentage of Non-Voters and Year of Birth (2016)"
+
+#plot did not vote
+plot_did_not_vote <- ggplot(percentage_data_long %>% filter(Candidate == "DidNotVote"), aes(x = birthyr, y = Percentage)) +
+  geom_point() + 
+  geom_smooth(method = "lm", se = FALSE, color = "orange") +
+  labs(title = "Percentage for Did Not Vote", x = "Birth Year", y = "Percentage (%)") +
+  theme_minimal()
+plot_did_not_vote
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-biden2020
+#| fig-cap: "Linear Regression comparing Percentage of Votes For Joe Biden and Year of Birth (2020)"
+
+ces2020_votes <- ces2020 %>%
+  filter(voted_for %in% c("Biden", "Trump"))
+
+# Calculate percentages for all categories
+percentage_data_2 <- ces2020_votes %>%
+  group_by(birthyr) %>%
+  summarise(
+    Biden = sum(voted_for == "Biden") / n() * 100,
+    Trump = sum(voted_for == "Trump") / n() * 100
+  ) %>%
+  ungroup()
+
+# Reshape data to long format
+percentage_data_2_long <- pivot_longer(percentage_data_2, cols = -birthyr, names_to = "Candidate", values_to = "Percentage")
+
+# Plot for Joe Biden
+plot_biden <- ggplot(percentage_data_2_long %>% filter(Candidate == "Biden"), aes(x = birthyr, y = Percentage)) +
+  geom_point() + 
+  geom_smooth(method = "lm", se = FALSE, color = "blue") +
+  labs(title = "Percentage for Joe Biden", y = "Percentage (%)", x = "Birth Year") +
+  theme_minimal()
+
+plot_biden
 
 ```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-trump2020
+#| fig-cap: "Linear Regression comparing Percentage of Votes For Donald Trump Year of Birth (2020)"
+
+# Plot for Donald Trump
+plot_trump_2 <- ggplot(percentage_data_2_long %>% filter(Candidate == "Trump"), aes(x = birthyr, y = Percentage)) +
+  geom_point() + 
+  geom_smooth(method = "lm", se = FALSE, color = "red") +
+  labs(title = "Percentage for Donald Trump", y = "Percentage (%)", x = "Birth Year") +
+  theme_minimal()
+
+plot_trump_2
+```
+
 ## Model 2

### 2024-03-16 paper/paper.qmd
added the weakness and stuff

--- 2024-03-16 paper/paper.qmd
@@
 
-# Models
+# Models/Results
 
@@
 
-# Results
-
 # Discussion
@@
 
-Weaknesses and next steps should also be included.
+  It is near impossible to have data sets that are completely error-free and the same is true for this data-set.
+The data-set we used as analysis does not take into account the opinions of non-citizens and who they would have voted for.
+A lot of marginalized opinions are being left out when we clean the data and we only look for people that could vote.
+For example refugees may have a different opinion on who got elected in the 2020 presidential candidate as they come from a different aspect of life and have different priorities that need to be addressed such as housing and or health care.
+  Another weakness in our data-set is what occurs with a lot of data-sets, which is missing values.
+This data set also contained them and those needed to be cleaned so that our analysis would not sway.
+Speaking of missing values, insignificant values were ignored in some of the models as well.
+For, example some of the presidential candidates got so little votes that including them would have actually hindered our ability to see the trends so we had to group them in others.
+This was beneficial for this paper but  if we do indeed continue to analyse this paper for other purposes, this might cause problems.
+
+  In terms of next steps for this paper, we look forward to predict the 2024 election results by comparing the what the general priorities of the population is.
+We could look at population by immigration status, gender, age, or their previous 2020 voting results.
+If we are able to correctly predict the results based on the model continuously, it would help understand the priorities of american citizens; knowledge that would in turn help us help them.
 

### 2024-03-17 paper/paper.qmd
added the posterior checks and rhat stuff to appendix

@@
 ```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-modelresults
+#| tbl-cap: "Explanatory models of presidential election in 2020 based on gender and immigration status"
+#| warning: false
 modelsummary(
@@
 
+pp_check(political_preferences) +
+  theme_classic() +
+  theme(legend.position = "bottom")
+
+posterior_vs_prior(political_preferences) +
+  theme_minimal() +
+  scale_color_brewer(palette = "Set1") +
+  theme(legend.position = "bottom") +
+  coord_flip()
+
 ```
@@
 
+plot(political_preferences, "trace")
+
+plot(political_preferences, "rhat")
+
 ```

### 2024-03-17 paper/paper.qmd
.

--- 2024-03-16 paper/paper.qmd
@@
 
-# Models/Results
+# Models
 
@@
 
-### Voters in 2016
+### Aging Voters (2016)
+
+In the following graphs we will be looking at the percentage of voters from each birth year that voted for Hillary Clinton, Donald Trump,and other candidates in the 2016 US Presidential election, we will also take a look at those who did not vote.
 
@@
 #| label: fig-hillary2016
-#| fig-cap: "Linear Regression comparing Percentage of Votes For Hillary Clinton and Year of Birth (2016)"
+#| fig-cap: "Comparing Percentage of Votes For Hillary Clinton and Year of Birth (2016)"
 
-# only choose values "Hillary Clinton" and "Donald Trump"
 ces2020_votes <- ces2020 %>%
@@
 
+#R value
+hillary_data <- percentage_data_long %>% filter(Candidate == "HillaryClinton")
+
+R_hillary <- cor(hillary_data$birthyr, hillary_data$Percentage)
+
 plot_hillary <- ggplot(percentage_data_long %>% filter(Candidate == "HillaryClinton"), aes(x = birthyr, y = Percentage)) +
@@
   geom_smooth(method = "lm", se = FALSE, color = "blue") +
-  labs(title = "Percentage for Hillary Clinton", y = "Percentage (%)") +
+  labs(title = "Percentage for Hillary Clinton", y = "Percentage (%)", x = "Birth Year") +
   theme_minimal()
@@
 ```
+
+In @fig-hillary2016 we can see that the younger the voters, the smaller percentage of their birth year seem to have voted for Hillary Clinton in the 2016 presidential election.
+This was rather surprising for us, as we expected young voters more likely to vote for Hillary Clinton and the Democratic Party, as younger people are often more progressive than their older counterparts.
+@gig-hillary shows us that there seems to be a significant correlation between age, birth year, and what percent of that birth year voted for Hillary with a correlation coefficient of $R = -0.94$, which is a rather strong negative relationship between our two variables.
 
@@
 #| label: fig-trump2016
-#| fig-cap: "Linear Regression comparing Percentage of Votes For Donald Trump and Year of Birth (2016)"
+#| fig-cap: "Comparing Percentage of Votes For Donald Trump and Year of Birth (2016)"
+
+trump_data <- percentage_data_long %>% filter(Candidate == "DonaldTrump")
+
+R_trump <- cor(trump_data$birthyr, trump_data$Percentage)
 
@@
   geom_smooth(method = "lm", se = FALSE, color = "red") +
-  labs(title = "Percentage for Donald Trump", y = "Percentage (%)") +
+  labs(title = "Percentage for Donald Trump", y = "Percentage (%)", x = "Birth Year") +
   theme_minimal()
@@
 ```
+
+While we look at the graph for Donald Trump in the 2016 election at first glance it does seem that the younger you are the more likey you were to have voted for trump in the 2016 election, but if we dig a bit deeper we can see that there seem to be several outliers on this graph, and the data points seem to somewhat be bouncing around a bit.
+This is confirmed when we look at our correlation coefficient $ R = 0.12$, so we can come to a conclsion that voting for Donald Trump in 2016 has a relatively weak positive correlation.
 
@@
 #| label: fig-other2016
-#| fig-cap: "Linear Regression comparing Percentage of Votes For Other Caniddates and Year of Birth (2016)"
+#| fig-cap: "Comparing Percentage of Votes For Other Caniddates and Year of Birth (2016)"
+
+other_data <- percentage_data_long %>% filter(Candidate == "Other")
+
+R_other <- cor(other_data$birthyr, other_data$Percentage)
 
@@
 
+@fig-other2016 does not seem to give us to much information about the age of those who voted for other candidates in the 2016 presidential elections, such as Gary Johnson and Jill Stein.
+While all these candidates received a combined less than 5% of the total vote, and less than 5% of the share of the vote from each birth year, we believe that transparency is incredibly important when doing data heavy research and while this may not seem as important as the graphs for Hillary Clinton and Donald Trump, this graph absolutely has a place in this paper.
+
 ```{r}
@@
 #| label: fig-dnv2016
-#| fig-cap: "Linear Regression comparing Percentage of Non-Voters and Year of Birth (2016)"
+#| fig-cap: "Comparing Percentage of Non-Voters and Year of Birth (2016)"
+
+dnv_data <- percentage_data_long %>% filter(Candidate == "Did not vote")
+
+R_dnv <- cor(dnv_data$birthyr, dnv_data$Percentage)
 
@@
 
+Here we will look at the members of the population that did not vote. @fig-dnv2016
+
 ```{r}
@@
 #| label: fig-biden2020
-#| fig-cap: "Linear Regression comparing Percentage of Votes For Joe Biden and Year of Birth (2020)"
+#| fig-cap: "Comparing Percentage of Votes For Joe Biden and Year of Birth (2020)"
 
@@
 
+#Rvalue Biden
+biden_data <- percentage_data_long %>% filter(Candidate == "Biden")
+
+R_biden <- cor(biden_data$birthyr, biden_data$Percentage)
+
 ```
@@
 #| label: fig-trump2020
-#| fig-cap: "Linear Regression comparing Percentage of Votes For Donald Trump Year of Birth (2020)"
+#| fig-cap: "Comparing Percentage of Votes For Donald Trump Year of Birth (2020)"
 
@@
 plot_trump_2
+
+#rvalue Trump 2020
+
+trump_2_data <- percentage_data_long %>% filter(Candidate == "Trump")
+
+R_trump_2 <- cor(trump_2_data$birthyr, trump_2_data$Percentage)
 ```
@@
 
+# Results
+
 # Discussion
@@
 
-  It is near impossible to have data sets that are completely error-free and the same is true for this data-set.
-The data-set we used as analysis does not take into account the opinions of non-citizens and who they would have voted for.
-A lot of marginalized opinions are being left out when we clean the data and we only look for people that could vote.
-For example refugees may have a different opinion on who got elected in the 2020 presidential candidate as they come from a different aspect of life and have different priorities that need to be addressed such as housing and or health care.
-  Another weakness in our data-set is what occurs with a lot of data-sets, which is missing values.
-This data set also contained them and those needed to be cleaned so that our analysis would not sway.
-Speaking of missing values, insignificant values were ignored in some of the models as well.
-For, example some of the presidential candidates got so little votes that including them would have actually hindered our ability to see the trends so we had to group them in others.
-This was beneficial for this paper but  if we do indeed continue to analyse this paper for other purposes, this might cause problems.
-
-  In terms of next steps for this paper, we look forward to predict the 2024 election results by comparing the what the general priorities of the population is.
-We could look at population by immigration status, gender, age, or their previous 2020 voting results.
-If we are able to correctly predict the results based on the model continuously, it would help understand the priorities of american citizens; knowledge that would in turn help us help them.
+Weaknesses and next steps should also be included.
 

### 2024-03-17 paper/paper.qmd

@@
 ```
-
-# Results
 

### 2024-03-17 paper/paper.qmd
.

@@
 This was rather surprising for us, as we expected young voters more likely to vote for Hillary Clinton and the Democratic Party, as younger people are often more progressive than their older counterparts.
-@gig-hillary shows us that there seems to be a significant correlation between age, birth year, and what percent of that birth year voted for Hillary with a correlation coefficient of $R = -0.94$, which is a rather strong negative relationship between our two variables.
+@fig-hillary shows us that there seems to be a significant correlation between age, birth year, and what percent of that birth year voted for Hillary with a correlation coefficient of $R = -0.94$, which is a rather strong negative relationship between our two variables.
 

### 2024-03-17 paper/paper.qmd
.

@@
 While we look at the graph for Donald Trump in the 2016 election at first glance it does seem that the younger you are the more likey you were to have voted for trump in the 2016 election, but if we dig a bit deeper we can see that there seem to be several outliers on this graph, and the data points seem to somewhat be bouncing around a bit.
-This is confirmed when we look at our correlation coefficient $ R = 0.12$, so we can come to a conclsion that voting for Donald Trump in 2016 has a relatively weak positive correlation.
+This is confirmed when we look at our correlation coefficient $R = 0.12$, so we can come to a conclsion that voting for Donald Trump in 2016 has a relatively weak positive correlation.
 
@@
 
-Here we will look at the members of the population that did not vote. @fig-dnv2016
+Here we will look at the members of the population that did not vote. @fig-dnv2016 shows us the steady increase of those citizens who do not vot
 

### 2024-03-17 paper/paper.qmd
.

@@
 
-Here we will look at the members of the population that did not vote. @fig-dnv2016 shows us the steady increase of those citizens who do not vot
+Here we will look at the members of the population that did not vote.
+@fig-dnv2016 shows us the steady increase of those citizens who did not vote in the 2016 presidential elections.
+We can see that they younger generation seems to have not voted for any candidate in the 2016 elections.
+This can be due to a variety of reasons, such as disengagement with the political process, dissatisfaction with the candidates presented, or a feeling that their vote may not make a significant difference in the outcome.
+Other contributing factors could include obstacles to voting access, such as registration issues, polling place locations, or work and school schedules that conflict with voting hours.
+
+As we further dissect these trends, it's essential to understand the broader societal and cultural influences that might be contributing to this detachment.
+
+### Aging Voters (2020)
+
+The data provided for the 2020 elections was slightly different the the data from the 2016 elections, the main difference was that, there was no one who voted for another candidate, other than Joe Biden and Donald Trump from the Democratic and Republican parties respectively.
+There were also no "Did not vote" inputs like there were in the data from the 2016 election between Donald and Hillary.
 
@@
 
+In @fig-biden2020 we see a complete opposite graph compared to Hillary's from 2016 where her's seemed to be more popular among older voters where as Biden seems to be much more popular among younger voters.
+So much so that voters born after 1980 had at least 60% of their age groups voting for Biden and those born after 1990 had at least 70% of their age groups voting for Biden.
+This is a much less surprising Graph than Hillary's, as younger voters seem to be more likely to vote for more progressive parties and candidates.
+
+We can see that for the most part the graph seems to be increasing at nearly every point except for a dip starting around 1950 and does not start to pick up until 1960, which seems somewhat odd until we recall that those are the prime birth year of the baby boomer generation.
+
 ```{r}
@@
 ```
+
+Lastly, we will look at @fig-trump2020 which is essentially the invers of the previous graph as there were no other possible data points other than "Biden" and "Trump"
 

### 2024-03-17 paper/paper.qmd
finsihed analysis and discussion for model 3

--- 2024-03-17 paper/paper.qmd
@@
 The 2024 elections is also coming up this year meaning that analyzing this results may help us predict what the outcome of the election might be.
+As for the results of the election in 2020, Donald Trump won the precidential title.
 
@@
 
-### Aging Voters (2016)
-
-In the following graphs we will be looking at the percentage of voters from each birth year that voted for Hillary Clinton, Donald Trump,and other candidates in the 2016 US Presidential election, we will also take a look at those who did not vote.
+### Voters in 2016
 
@@
 #| label: fig-hillary2016
-#| fig-cap: "Comparing Percentage of Votes For Hillary Clinton and Year of Birth (2016)"
+#| fig-cap: "Linear Regression comparing Percentage of Votes For Hillary Clinton and Year of Birth (2016)"
 
+# only choose values "Hillary Clinton" and "Donald Trump"
 ces2020_votes <- ces2020 %>%
@@
 
-#R value
-hillary_data <- percentage_data_long %>% filter(Candidate == "HillaryClinton")
-
-R_hillary <- cor(hillary_data$birthyr, hillary_data$Percentage)
-
 plot_hillary <- ggplot(percentage_data_long %>% filter(Candidate == "HillaryClinton"), aes(x = birthyr, y = Percentage)) +
@@
   geom_smooth(method = "lm", se = FALSE, color = "blue") +
-  labs(title = "Percentage for Hillary Clinton", y = "Percentage (%)", x = "Birth Year") +
+  labs(title = "Percentage for Hillary Clinton", y = "Percentage (%)") +
   theme_minimal()
@@
 ```
-
-In @fig-hillary2016 we can see that the younger the voters, the smaller percentage of their birth year seem to have voted for Hillary Clinton in the 2016 presidential election.
-This was rather surprising for us, as we expected young voters more likely to vote for Hillary Clinton and the Democratic Party, as younger people are often more progressive than their older counterparts.
-@fig-hillary shows us that there seems to be a significant correlation between age, birth year, and what percent of that birth year voted for Hillary with a correlation coefficient of $R = -0.94$, which is a rather strong negative relationship between our two variables.
 
@@
 #| label: fig-trump2016
-#| fig-cap: "Comparing Percentage of Votes For Donald Trump and Year of Birth (2016)"
-
-trump_data <- percentage_data_long %>% filter(Candidate == "DonaldTrump")
-
-R_trump <- cor(trump_data$birthyr, trump_data$Percentage)
+#| fig-cap: "Linear Regression comparing Percentage of Votes For Donald Trump and Year of Birth (2016)"
 
@@
   geom_smooth(method = "lm", se = FALSE, color = "red") +
-  labs(title = "Percentage for Donald Trump", y = "Percentage (%)", x = "Birth Year") +
+  labs(title = "Percentage for Donald Trump", y = "Percentage (%)") +
   theme_minimal()
@@
 ```
-
-While we look at the graph for Donald Trump in the 2016 election at first glance it does seem that the younger you are the more likey you were to have voted for trump in the 2016 election, but if we dig a bit deeper we can see that there seem to be several outliers on this graph, and the data points seem to somewhat be bouncing around a bit.
-This is confirmed when we look at our correlation coefficient $ R = 0.12$, so we can come to a conclsion that voting for Donald Trump in 2016 has a relatively weak positive correlation.
 
@@
 #| label: fig-other2016
-#| fig-cap: "Comparing Percentage of Votes For Other Caniddates and Year of Birth (2016)"
-
-other_data <- percentage_data_long %>% filter(Candidate == "Other")
-
-R_other <- cor(other_data$birthyr, other_data$Percentage)
+#| fig-cap: "Linear Regression comparing Percentage of Votes For Other Caniddates and Year of Birth (2016)"
 
@@
 
-@fig-other2016 does not seem to give us to much information about the age of those who voted for other candidates in the 2016 presidential elections, such as Gary Johnson and Jill Stein.
-While all these candidates received a combined less than 5% of the total vote, and less than 5% of the share of the vote from each birth year, we believe that transparency is incredibly important when doing data heavy research and while this may not seem as important as the graphs for Hillary Clinton and Donald Trump, this graph absolutely has a place in this paper.
-
 ```{r}
@@
 #| label: fig-dnv2016
-#| fig-cap: "Comparing Percentage of Non-Voters and Year of Birth (2016)"
-
-dnv_data <- percentage_data_long %>% filter(Candidate == "Did not vote")
-
-R_dnv <- cor(dnv_data$birthyr, dnv_data$Percentage)
+#| fig-cap: "Linear Regression comparing Percentage of Non-Voters and Year of Birth (2016)"
 
@@
 
-Here we will look at the members of the population that did not vote. @fig-dnv2016
-
 ```{r}
@@
 #| label: fig-biden2020
-#| fig-cap: "Comparing Percentage of Votes For Joe Biden and Year of Birth (2020)"
+#| fig-cap: "Linear Regression comparing Percentage of Votes For Joe Biden and Year of Birth (2020)"
 
@@
 
-#Rvalue Biden
-biden_data <- percentage_data_long %>% filter(Candidate == "Biden")
-
-R_biden <- cor(biden_data$birthyr, biden_data$Percentage)
-
 ```
@@
 #| label: fig-trump2020
-#| fig-cap: "Comparing Percentage of Votes For Donald Trump Year of Birth (2020)"
+#| fig-cap: "Linear Regression comparing Percentage of Votes For Donald Trump Year of Birth (2020)"
 
@@
 plot_trump_2
-
-#rvalue Trump 2020
-
-trump_2_data <- percentage_data_long %>% filter(Candidate == "Trump")
-
-R_trump_2 <- cor(trump_2_data$birthyr, trump_2_data$Percentage)
 ```
@@
 y_i|\pi_i  &\sim \mbox{Bern}(\pi_i) \\
-\mbox{logit}(\pi_i) &= \beta_0 + \beta_1 \times \mbox{gender}_i + \beta_2\times \mbox{immigration status}_i\\
+\mbox{logit}(\pi_i) &= \beta_0 + \beta_1 \times \mbox{gender}_i + \beta_2\times \mbox{immigration status}_i + \beta_3\times \mbox{previous 2016 vote}_i\\
 \beta_0 &\sim \mbox{Normal}(0, 2.5) \\
 \beta_1 &\sim \mbox{Normal}(0, 2.5) \\
-\beta_2 &\sim \mbox{Normal}(0, 2.5)
+\beta_2 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_3 &\sim \mbox{Normal}(0, 2.5)
 \end{aligned}
@@
   stan_glm(
-    voted_for ~ gender + immstat,
+    voted_for ~ gender + immstat + presvote16post ,
     data = ces2020_reduced,
@@
 #| eval: true
-#| label: tbl-modelresults
+#| label: tbl-modelresults3
 #| tbl-cap: "Explanatory models of presidential election in 2020 based on gender and immigration status"
@@
 
+The numbers on the first half of the table @tbl-modelresults3, right next to each variable name, shows us the coefficient related to each of the predictor variables.
+It tells us whether the respondent is expected to vote for Biden or not based on their immigration status, gender and who they voted for in the 2016 elections.
+The bracketed values are the errors associated with the coefficients so the smaller the number the less likely the model inaccurately came up with the coefficients.
+Notice that one of the variables is named intercept which essentially represents gender Female.
+We see that in "parent and I born in US but grandparent(s) immigrant", "gender male" and "gender female", the coefficients are negative meaning that they are not inclined to vote for Biden in the 2020 election.
+All other values are positive and we see for people who voted for Trump in the 2016 election, they have a higher coefficient with a relatively small inaccuracy rate.
+
 # Discussion
@@
 
-## Third discussion point
+## Modeling immigration status, gender and 2016 voting results
+
+   @tbl-modelresults3 was a very interesting plot to examine as they let us see the correlation between the election held in 2020 and other factors such as gender, immigration status, and the election results in 2016.
+Starting off we look at the negative values.
+We see that both males and females are less likely to vote for Biden as their coefficient is negative.
+The female has more of a negative value meaning that their priorities do not align with the view of the democratic party.
+This makes sense as we look at the news in the past years when we see the Biden government push for anti-abortion laws, the very law that a lot of women are opposed to as they argue they deserve to make that choice.
+As for the negative value for the males, there does not seem to be a direct answer to that.
+From here we see that it makes more sense to look at the large coefficients as the smaller values may be a result of many different things.
+A person is just not male but they may also be from a different ethnicity or religion whose priorities are different from the Biden government.
+America is a very diverse country so it makes sense that many people have different reasons to support and not support the republicans.
+The next largest value we notice is for people who were born in the US but have immigrant parents and people who are immigrants but not citizens.
+This makes sense as the Biden government actually reversed some of the immigration plans implemented by the previous government that harmed immigrants.
+Some of these laws were the travel ban put on Muslim-majority countries that resulted in many immigrants being unable to come to America to find a better future.
+It makes sense that children of people who immigrated and people who are immigrants themselves would prioritize such issues.
+Lastly, we look at the another coefficient found in the table which is the people who voted for Trump in the 2016 election.
+We see that it is very likely that people who voted for Trump in the last election, voted for Biden in this one.
+This may be due to the fact that there were many protests of the Trump administration and how they ran the country during their term.
+Some of the backlash that happened while Trump was president was the travel ban on Muslim-majority countries, reduced government regulation across various industries, and tax cuts from private corporations.
+These decisions may have led many people to not re-elect Trump for another 4 years as their president.
+The same is seen for people who voted for Evan McMullin (independent candidate) and other smaller parties.
+The reasons are hard to infer but it may be due to the fact that their candidate did not run again or their priorities and views changed from the party/person they voted for.
 
@@
 
-  It is near impossible to have data sets that are completely error-free and the same is true for this data-set.
-The data-set we used as analysis does not take into account the opinions of non-citizens and who they would have voted for.
+  It is nearly impossible to have data sets that are completely error-free and the same is true for this data set.
+The data set we used as analysis does not take into account the opinions of non-citizens and who they would have voted for.
 A lot of marginalized opinions are being left out when we clean the data and we only look for people that could vote.
-For example refugees may have a different opinion on who got elected in the 2020 presidential candidate as they come from a different aspect of life and have different priorities that need to be addressed such as housing and or health care.
-  Another weakness in our data-set is what occurs with a lot of data-sets, which is missing values.
+For example, refugees may have a different opinion on who got elected in the 2020 presidential candidate as they come from a different aspect of life and have different priorities that need to be addressed such as housing and or health care.
+  Another weakness in our data set is what occurs with a lot of data sets, which is missing values.
 This data set also contained them and those needed to be cleaned so that our analysis would not sway.
 Speaking of missing values, insignificant values were ignored in some of the models as well.
-For, example some of the presidential candidates got so little votes that including them would have actually hindered our ability to see the trends so we had to group them in others.
-This was beneficial for this paper but  if we do indeed continue to analyse this paper for other purposes, this might cause problems.
+For, example some of the presidential candidates got so few votes that including them would have actually hindered our ability to see the trends so we had to group them into others.
+This was beneficial for this paper but if we do indeed continue to analyze this paper for other purposes, this might cause problems.
 
-  In terms of next steps for this paper, we look forward to predict the 2024 election results by comparing the what the general priorities of the population is.
-We could look at population by immigration status, gender, age, or their previous 2020 voting results.
-If we are able to correctly predict the results based on the model continuously, it would help understand the priorities of american citizens; knowledge that would in turn help us help them.
+  In terms of the next steps for this paper, we look forward to predicting the 2024 election results by comparing the general priorities of the population.
+We could look at the population by immigration status, gender, age, or their previous 2020 voting results.
+If we are able to correctly predict the results based on the model continuously, it would help us understand the priorities of American citizens; knowledge that would in turn help us help them.
 
@@
 
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-modelCred
+#| fig-cap: "Credible intervals for predictors of support for Biden"
+#| warning: false
+
+modelplot(political_preferences, conf_level = 0.9) +
+  labs(x = "90 per cent credibility interval")
+```
+
 # Model details {#sec-model-details}
@@
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
+In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check.
 
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior.
 

### 2024-03-17 paper/paper.qmd
.

--- 2024-03-17 paper/paper.qmd
@@
 
-Lastly, we will look at @fig-trump2020 which is essentially the invers of the previous graph as there were no other possible data points other than "Biden" and "Trump"
+In @fig-biden2020 we see a complete opposite graph compared to Hillary's from 2016 where her's seemed to be more popular among older voters where as Biden seems to be much more popular among younger voters.
+So much so that voters born after 1980 had at least 60% of their age groups voting for Biden and those born after 1990 had at least 70% of their age groups voting for Biden.
+This is a much less surprising Graph than Hillary's, as younger voters seem to be more likely to vote for more progressive parties and candidates.
+
+We can see that for the most part the graph seems to be increasing at nearly every point except for a dip starting around 1950 and does not start to pick up until 1960, which seems somewhat odd until we recall that those are the prime birth year of the baby boomer generation.
+
+Lastly, we will look at @fig-trump2020 which is essentially the inverse of the previous graph as there were no other possible data points other than "Biden" and "Trump".
+While this graph may just be a reflection across the 50% line, it is still helpful so that wer may look at it from a different perspective.
+Just as in @fig-biden2020, @fig-trump2020 has a small peak after the year 1950.
+It is also important to note that even though there is a peak after the year 1950, all of the ages that had more than a 50% vote share for trump all also came before the year 1950.
 

### 2024-03-17 paper/paper.qmd
.
Added subheaders

@@
 
-## Aging and Its Impact on Voting
+## The Generational Vote
 
@@
 
-### Aging Voters (2016)
+### The Generational Vote (2016)
 
@@
 
-### Aging Voters (2020)
+### The Generational Vote (2020)
 
@@
 ```
-
-In @fig-biden2020 we see a complete opposite graph compared to Hillary's from 2016 where her's seemed to be more popular among older voters where as Biden seems to be much more popular among younger voters.
-So much so that voters born after 1980 had at least 60% of their age groups voting for Biden and those born after 1990 had at least 70% of their age groups voting for Biden.
-This is a much less surprising Graph than Hillary's, as younger voters seem to be more likely to vote for more progressive parties and candidates.
-
-We can see that for the most part the graph seems to be increasing at nearly every point except for a dip starting around 1950 and does not start to pick up until 1960, which seems somewhat odd until we recall that those are the prime birth year of the baby boomer generation.
 
@@
 
-## First discussion point {#sec-first-point}
+## Dissecting Age Related Demographics
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+Our exploration into the voting patterns across different age demographics has surfaced intriguing insights into the electoral pulse of America during the 2016 and 2020 presidential elections.
+In 2016, the expectation was that Hillary Clinton would enjoy robust support from younger voters, yet the data depicted in @fig-hillary2016 revealed a contrary trend: a decrease in support among this demographic.
+The analysis of the voting trends across the 2016 and 2020 elections has brought to light notable differences in age related voting behaviors, especially concerning Donald Trump's candidacy.
+In the 2016 election, as depicted by @fig-trump2016, the expectation of a clear trend regarding age and support for Trump was not met.
+Instead, we observed a mildly chaotic scatter of data points, indicating no strong age-based preference.
+This was quantitatively supported by a weak positive correlation coefficient, R = 0.12, suggesting that Trump's voter support did not significantly increase or decrease with voters' ages.
+Alternately, in 2020, the political landscape has shifted with no small part to the global pandemic COVID-19, public discourse and other political movements including but not limited to the death of George Floyd.
+After many events in Trump's presidency, some within his control and some not, we have the information from @fig-trump2020, where we see more of as pattern among the data points where we can see Trump is more favored among older generations.
+This suggests an intensification of partisan divides along generational lines, with older voters becoming more entrenched in their support for Trump, while younger voters showed increasing support for the opposing Democratic candidate, Joe Biden.
 
@@
 This was beneficial for this paper but if we do indeed continue to analyze this paper for other purposes, this might cause problems.
+  Additionally, in the 2016 data there are more candidates that the people voted fore, while not receiving many votes, there were absolutely other candidates mentioned in the 2020 data, as well as no people from the surveyed population chose not to vot unlike in the 2016 election, which could have provided a deeper insight on the differences between the two elections.
 

### 2024-03-17 paper/paper.qmd
minor edits
Update paper.qmd

@@
 
-## Model 2
+## Comparing 2016 and 2020 votes based on birth year and gender
 In this model, we conduct an examination of the relationship between voters' birth year and their gender for the 2016 and 2020 Presidential elections.
@@
 
-kables(list(votes_2020_f, votes_2020_m))
+votes_2020_f
+votes_2020_m
 #If the ratio is greater than 1, it means that within the group, Biden has more votes than Trump
@@
 
-kables(list(votes_2020_f, votes_2020_m))
-#If the ratio is greater than 1, it means that within the group, Biden has more votes than Trump
-#if the ratio is less than 1, Trump has more votes within the group
+votes_2020_f
+votes_2020_m
 ```
@@
 
-## Model 3
+## Predicitng 2020 votes based off regression variables
 
@@
 
-## Dissecting Age Related Demographics
+## Dissecting Age Related Voting Patterns
 
@@
 ## Second discussion point
+
+## Comparing 2016 and 2020 votes votes based on birth year and gender
+
+Our analysis displayed interesting results when when examining gender-based voting behaviors the from 2016 and 2020 US Presidential election votes in @fig-genderyr2016 and @fig-genderyr2020.
+We notice from the graphs and in @tbl-ratio2020 most votes were for the Republican Party in 2016, while most votes for the Democratic Party in 2020.
+Female voters have a higher ratio of Democratic voters to Republican voters in both years.
+In contrast, the distribution of male votes between the two parties was more evenly balanced.
+The most interesting and abstract result is that the votes are reversed in 2016 compared to 2020.
+There can be many explanations for this reverse, including that Americans were not satisfied with the Republican Presidency from 2016 to 2020.
+Further research should be done into the reason why there is a higher ratio of Democratic to Republican female voters compared to male voters.
+
+## Third discussion point
 
